---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# leninature

<!-- badges: start -->
<!-- badges: end -->

The goal of this package is to make it easy to query and analyse the
works of Lenin. 

# Getting started

If you would like to use the tidy data in the format provided, simply install
this package from github using devtools:

```{r eval=FALSE}
devtools::install_github("alicirce/leninature")
```

If you would like to run the data compilation scripts yourself from scratch, you
will need to download the texts from the Marxists Internet Archive (MIA). MIA
provides instructions for (respectfully) downloading portions of their archives
in their [FAQ](https://www.marxists.org/admin/janitor/faq.htm#hdd).

The scripts in `data-raw` assume you have already run the following scripts. 
You will also need to have your working directory appropriately set.

```{bash eval=FALSE}
mkdir lenin
cd lenin
wget -mpnp -nH -N -t 3 -w .5 https://www.marxists.org/archive/lenin/by-title.htm
```

This took me about 2.5 hours to pull.
