---
title: "Literacy"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{literacy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  echo = FALSE,
  fig.align = "center"
)
```

```{css}
img {
  border: 0px;
}
p.caption {
  font-size: 0.8em;
}
```

```{r setup}
library(leninature)
library(dplyr)
library(ggplot2)
```

> As regards the second enemy, illiteracy, I can say that so long as there is
such a thing as illiteracy in our country it is too much to talk about political
education. This is not a political problem; it is a condition without which it
is useless talking about politics. An illiterate person stands outside politics,
he must first learn his ABC. Without that there can be no politics; without that
there are rumours, gossip, fairy-tales and prejudices, but not politics.
- [Lenin (1921)](https://www.marxists.org/archive/lenin/works/1921/oct/17.htm)

Socialist governments tend to radically increase literacy rates. In a little 
over a year, [Cuba raised literacy rates from 47% to 100%](https://daily.jstor.org/rosa-hernandez-acosta-on-the-cuban-literacy-campaign/).
China raised literacy from [20% in 1950 to more than 85% in 2001](nytimes.com/2001/02/12/news/chinas-long-but-uneven-march-to-literacy.html).
Lenin's administration similarly viewed it as a top priority, naming illiteracy
as one of three enemies in his 1921 address on the New Economic Policy.

However, revolutionary efforts to improve literacy not only predate these socialist
revolutions, but paved their way. Bolsheviks and other revolutionary parties
realized that one of the primary barriers to raising the political consciousness
of the workers was their illiteracy. In 1897, only about 
[1 in 3 men and 1 in 10 women](https://www.marxists.org/archive/lenin/works/1897/handicraft/2iv.htm)
were literate. [In 1913](https://www.marxists.org/archive/lenin/works/1913/apr/27c.htm), 
Lenin referred to the literacy rate of 27% as "stultifying" the people.
Revolutionaries  (including Krupskaya) taught in 
[Sunday schools](https://twitter.com/alicirce/status/1553175614093266944?s=20&t=A0KnlZurDZUJpUVPjQWU0Q), 
geared towards workers on their days off, and these venues became places for
raising literacy rates as well as political consciousness. 

You can see this consistent emphasis on literacy in Lenin's writing throughout
the years, both before the 1917 October Revolution, as part of political
education of the working class, and after the revolution, as the governing
party.

```{r fig.height=3,fig.width=4}
lenin_literacy <- lenin %>%
  filter(grepl("literacy|literate", text, ignore.case = T)) %>%
  leninature:::count_mentions_in_dataframe(
    words = c(
      "literacy", "illiteracy", 
      "illiterate", "literate",
      "illiterates", "literates"
    )
  ) 

lit_by_year <- lenin_literacy %>%
  mutate(word = ifelse(
    grepl("illit", word),
    "illiterate/illiteracy",
    "literate/literacy"
  )) %>%
  group_by(word, year) %>%
  summarize(n = sum(mentions))

lit_by_year %>%
  ggplot() +
  aes(x = year, y = n) +
  geom_hline(yintercept = 0, color = "grey40") +
  geom_bar( 
    stat = "identity",
    #fill = "grey60",
    aes(fill = word),
    width = 0.5
  ) +
  theme_void() +
  theme(
    legend.position = 'bottom',
    legend.key.height = unit(0.2, "pt"),
    legend.title = element_blank(),
    strip.background = element_blank(),
    axis.text = element_text(color = 'grey30', size = 8),
    axis.ticks.x = element_line(color = "grey60"),
    axis.ticks.length = unit(0.1, units = "in"),
    panel.grid.major.y = element_line(color = "grey90"),
    plot.background = element_rect(fill = "white", color = NA),
    plot.title = element_text(size = 10, color = "#f20707", face = "bold"),
    plot.subtitle = element_text(size = 9, color = "grey30"),
    plot.caption = element_text(size = 7, color = "grey30"),
    strip.text = element_blank()
  ) +
  labs(
    title = "Lenin & literacy",
    subtitle = "Number of mentions of literacy/illiteracy"
  ) +
  scale_y_continuous(expand = c(0.0, 0., 0.2, 0.2)) +
  scale_x_continuous(breaks = seq(1895, 1920, 5)) +
  scale_fill_manual(
    values = c("grey60", "grey40")
  )
```

What were they reading in these literacy groups and reading circles?
Chernychevsky's *What is to be done?*! One such participant, who later joined
the Bolshevik party and participated in the 1917 Revolution reflected:

> When we started to attend the Sunday and evening schools, we began to make use
of books from the library and we learned of the great Russian democrat,
Chernychevsky. We read his book *What is to be done?*, secretly, and found the
image of Vera Pavolovna, the woman of the future, very attractive
- Alexandra Artiukhina, [quoted in Midwives of the Revolution](https://www.taylorfrancis.com/books/mono/10.4324/9780203981603/midwives-revolution-jane-mcdermid-anna-hillyar)


## Why is this important today?

If the argument against encouraging people sympathetic to socialist movements to
read or critically engage with socialist texts is that literacy in the general
public is too low (eg., 54% of the US adult population reading
[below a 6th grade level](https://www.forbes.com/sites/michaeltnietzel/2020/09/09/low-literacy-levels-among-us-adults-could-be-costing-the-economy-22-trillion-a-year/?sh=303aa1944c90)) then the 
solution is to improve literacy, and help people to learn and understand, not
reduce movement beliefs to bite-size slogans.
