---
title: "Mentions of Mencheviks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mentions of Mencheviks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  echo = FALSE,
  fig.align = "center"
)
```

```{css}
img {
  border: 0px;
}
p.caption {
  font-size: 0.8em;
}
```

```{r setup}
library(leninature)
library(dplyr)
library(tidyr)
library(ggplot2)
```

A lot of Lenin's writing takes the form of arguments against political or
philosophical opponents. In [*The Proletarian Revolution and the Renegade Kautsky*](https://www.marxists.org/archive/lenin/works/1918/prrk/) (1918), Lenin 
argues against Social Democrat Kautsky about the role of democracy and revolution 
in  the transition to socialism. 
In [*The Economic Content of Narodism and the Criticism of it in Mr. Struve’s Book*](https://www.marxists.org/archive/lenin/works/1894/narodniks/index.htm),
dissects Peter Struve's 1894 book, accusing it of being a liberal bourgeois 
distortion of Marxism. I wondered who was most frequently mentioned in Lenin's
works. (See a 
[similar analysis](https://alicirce.github.io/marxmywords/articles/capital_footnotes.html#citation-content) 
I did for Marx's *Capital*)

## Methods

I used {spacyr}'s parsing tools to tag parts of speech. This tool uses sentence
syntax structure to identify what words are verbs, nouns, adjectives, etc. I
then looked for the most common words (or series of words) tagged as proper
nouns. I manually classified the top hits as being associated with people,
political parties, government institutions, or newspapers.

```{r}

# The most frequently occurring proper nouns with at least 160 instances,
# manually grouped according to type of pronoun. This list excludes things
# labelled as proper nouns related to places (eg, Moscow) or titles (eg, Mr), 
# and roughly corresponds to the top ~270 proper nouns.
top_proper_nouns <- select_proper_nouns <- list(
  people = c(
    "Marx", "Engels", "Kautsky", "Trotsky", "Kerensky", "Mach", "Martov",
    "Sismondi", "Struve", "Kolchak", "Maslov", "Potresov", "Axelrod", 
    "Bogdanov", "Stolypin", "Milyukov", "Denikin", "Avenarius", "Bulgakov",
    "Chkheidze", "Martynov", "Hegel", "Larin", "Radek", "Bernstein",
    "Rosa Luxemburg", "Mikhailovsky", "Kant", "Chernov", "Kamenev", "Bebel",
    "Tsereteli", "Grimm", "Stalin", "Liebknecht", "Guchkov", "Nadya", "Parvus",
    "Zinoviev", "Kornilov"
  ),
  government = c(
    "Central Committee", "C.C.", "Zemstvo", "Soviets", "Constituent Assembly",
    "Duma", "Commissariat", "Central Organ", "Soviet", "Red Army", 
    "Politbureau", "Provisional Government"
  ),
  parties = c(
    "Bolsheviks", "Menshevik", "Narodniks", "Mensheviks", "Narodnik", 
    "Bolshevik", "Trudoviks", "Trudovik", "R.S.D.L.P.", "Marxists", "Marxist",
    "Communists", "Octobrists", "Democratic Labour Party", "Black Hundreds",
    "Cadets", "Bund", "Bundists", "Narodism", "Bolshevism", "Menshevism",
    "Socialists"
  ),
  newspapers = c(
    "Iskra", "Pravda", "Vperyod", "Rech", "Rabocheye Dyelo", "Nasha Zarya",
    "Osvobozhdeniye", "Novoye Vremya", "Russkoye Bogatstvo", "Proletary"
  )
)

# I've retrospectively limited the vectors of key words to most common ones when
# assessed by total mentions (instead of restricting counts to those instances
# recognized by Spacyr's parts-of-speech tagging) so that this vignette builds a
# little faster, but you can verify for yourself.

#newspapers <- count_lenin_mentions(top_proper_nouns$newspapers[1:9])
people <- count_lenin_mentions(top_proper_nouns$people[1:15])


# parties is a little harder because it's sometimes in plural, sometimes not.
# I considered using word stems instead, but that doesn't work so well when 
# the party name is multiple words. This is a little bit of a hacky solution
# but I think it's a reasonable way to compare mentions within the parties.

# lenin_with_adjustments <- lenin %>%
#   mutate(
#     text = gsub("\\bbolshevik\\b", "Bolsheviks", text, ignore.case = T),
#     text = gsub("\\bbolshevism\\b", "Bolsheviks", text, ignore.case = T),
#     text = gsub("\\bmenshevik\\b", "Mensheviks", text, ignore.case = T),
#     text = gsub("\\bmenshevism\\b", "Mensheviks", text, ignore.case = T),
#     text = gsub("\\bnarodnik\\b", "Narodniks", text, ignore.case = T),
#     text = gsub("\\bnarodism\\b", "Narodniks", text, ignore.case = T),
#     text = gsub("\\btrudovik\\b", "Trudoviks", text, ignore.case = T),
#     text = gsub("\\btrudovism\\b", "Trudoviks", text, ignore.case = T),
#     text = gsub("R.S.D.L.P.", "Social-Democrats", text, ignore.case = T),
#     text = gsub("RSDLP", "Social-Democrats", text, ignore.case = T),
#     text = gsub(
#       "Socialist-Revolution", 
#       "Socialist-Revolutionaries", 
#       text, 
#       ignore.case = T
#     ),
#     text = gsub(
#       "\\bSocialist Revolutionaries\\b", 
#       "Socialist-Revolutionaries",
#       text, 
#       ignore.case = T
#     ),
#     text = gsub("Social-Democracy", "Social-Democrats", text, ignore.case = T),
#     text = gsub(
#       "Social-Democratic\\b", 
#       "Social-Democrats", 
#       text, 
#       ignore.case = T
#     ),
#     text = gsub(
#       "\\bSocial Democrats\\b",
#        "Social-Democrats",
#        text,
#        ignore.case = T
#     )
#   )

# parties <- leninature:::count_mentions_in_dataframe(
#   c(
#     "Bolsheviks", "Mensheviks", "Narodniks", 
#     "Trudoviks", 
#     #"RSDLP", 
#     "Social-Democrats"
#   ),
#   lenin_with_adjustments
# )

```

## People

Unsurprisingly, Marx makes it to the top of the list, with Engels not too far
behind. Lenin's wife, Krupskaya, relates the following anecdote in her
[essay on how Marx shaped Lenin's thinking](https://www.marxists.org/archive/krupskaya/works/howleninstudiedmarx.htm):

> For Lenin the teachings of Marx were a guide to action. He once used the
following expression: "Who wants to consult with Marx ? " . . . It is a very
characteristic expression. He himself constantly "consulted with Marx."

```{r fig.width=4, fig.height=3}
top_n <- people %>%
  group_by(word) %>% 
  summarize(n = sum(mentions)) %>%
  arrange(desc(n)) %>%
  slice(1:9)

top_n %>%
  ggplot() +
  aes(x = n, y = reorder(word, n)) +
  geom_point() +
  geom_segment(aes(xend = 0, yend = word)) +
  geom_label(
    aes(x = 0, label = word),
    fill = "white",
    label.size = NA,
    label.padding = unit(0, "cm"),
    hjust = 0,
    nudge_x = 80,
    size = 3.2
  ) +
  theme_classic() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.ticks.x = element_line(color = "black"),
    axis.line.y = element_line(size = 0.4),
    axis.title.y = element_blank(),
    plot.background = element_rect(fill = "white", color = NA),
    plot.title = element_text(size = 10, color = "#f20707", face = "bold"),
    plot.subtitle = element_text(size = 9, color = "grey30")
  ) +
  scale_x_continuous(
    expand = c(0, 0, 0.01, 0.01)
  ) +
  labs(
    title = "Lenin: friends and foes",
    subtitle = "Number of mentions across all Lenin's works (1893-1923)",
    x = "number of mentions"
  )

```

Kautsky, as I mentioned above, was a Social Democrat and opponent of Bolshevism, 
and the target of many works by Lenin. Martov led the Mensheviks, the faction of
the Russian Social Democratic Labour Party (RSDLP) that opposed the Bolsheviks
during the 2nd Party Congress in 1903. Trotsky was a sometimes-opponent and 
[sometimes-supporter](https://redsails.org/trotsky-on-state-capitalism/) of 
Lenin. Struve was a prominent member of the RSDLP, but became more liberal in
his politics over time and opposed Lenin's Bolsheviks. Stolypin was the prime
minister of Russia from 1906-1911. Axelrod was an editor (along with Lenin) of
the newspaper *Iskra*, and a member of the Menshevik faction. Mach was a
scientist (of Mach number fame) and a philosopher.

We can also look at how these people were mentioned across Lenin's writing.

```{r fig.width=6, fig.height=4.5}


people_graph_df <- people %>%
  mutate(person = as.character(word)) %>%
  filter(person %in% top_n$word) %>%
  group_by(year, person) %>%
  summarize(n = sum(mentions))  %>%
  ungroup() %>%
  group_by(person) %>%
  mutate(tot = sum(n)) %>%
  ungroup() %>%
  mutate(person = reorder(person, desc(n)))

# separate
people_graph_df  %>%
  ggplot() +
  aes(x = year, y = n) +
  geom_hline(yintercept = 0, color = "grey40") +
  geom_bar( 
    stat = "identity",
    fill = "grey60",
    width = 0.4
  ) +
  geom_text(
    data = select(people_graph_df, person) %>% distinct(), 
    aes(label = person, x = 1922, y = 650), 
    hjust = 1, size = 3.5
  ) +
  theme_void() +
  facet_wrap(~person, scales = "free_x") +
  theme(
    legend.position = 'none',
    strip.background = element_blank(),
    axis.text = element_text(color = 'grey30', size = 8),
    axis.ticks.x = element_line(color = "grey60"),
    axis.ticks.length = unit(0.1, units = "in"),
    panel.grid.major.y = element_line(color = "grey90"),
    plot.background = element_rect(fill = "white", color = NA),
    plot.title = element_text(size = 10, color = "#f20707", face = "bold"),
    plot.subtitle = element_text(size = 9, color = "grey30"),
    strip.text = element_blank()
  ) +
  labs(
    title = "Lenin: friends and foes",
    subtitle = "Frequently mentioned people (name drops per year)"
  ) +
  scale_y_continuous(expand = c(0.0, 0., 0.1, 0.1)) +
  scale_x_continuous(breaks = seq(1895, 1920, 5))


```

Perhaps the most unique pattern is that of Mach; 519 of the total 544 mentions
of Mach were in [*Materialism and Empirio-Criticism*](https://www.marxists.org/archive/lenin/works/1908/mec/index.htm),
(1908) Lenin's [longest work](https://alicirce.github.io/leninature/articles/leninature.html#books), 
in which Lenin argues against Mach's claims that no knowledge exists outside of
one's own senses and experience. The mentions of Struve, Stolypin, Struve 
correspond with the timelines of their political prominence. Marx features
fairly consistently throughout Lenin's work. Kautsky and Engels also span the
whole timeline of Lenin's writing but with a few sharp peaks. The sharp spike in
Engel's timeline in the graph represents *Materialism and Empirio-Criticism*
(409 out of 1674 mentions). The two tallest peaks for Kautsky are the 
aforementioned [*The Proletarian Revolution and the Renegade Kautsky*](https://www.marxists.org/archive/lenin/works/1918/prrk/) and the 1899
pair of articles titled 
[*Capitalism in Agriculture*](https://www.marxists.org/archive/lenin/works/1899/agriculture/).



```{r eval=FALSE}

party_graph_df <- parties %>%
  mutate(party = gsub("-", "\n", as.character(word))) %>%
  group_by(year, party) %>%
  summarize(n = sum(mentions))  %>%
  ungroup() %>%
  group_by(party) %>%
  mutate(tot = sum(n)) %>%
  ungroup() %>%
  mutate(party = reorder(party, desc(tot)))


party_graph_df  %>%
  ggplot() +
  aes(x = year, y = n) +
  geom_hline(yintercept = 0, color = "grey40") +
  geom_bar( 
    stat = "identity",
    #fill = "grey60",
    aes(fill = year == 1917),
    width = 0.5
  ) +
  geom_text(
    data = select(party_graph_df, party) %>% distinct(), 
    aes(label = party, x = 1922, y = 1200), 
    hjust = 1, size = 3.3
  ) +
  theme_void() +
  facet_wrap(~party, scales = "free_x") +
  theme(
    legend.position = 'none',
    strip.background = element_blank(),
    axis.text = element_text(color = 'grey30', size = 8),
    axis.ticks.x = element_line(color = "grey60"),
    axis.ticks.length = unit(0.1, units = "in"),
    panel.grid.major.y = element_line(color = "grey90"),
    plot.background = element_rect(fill = "white", color = NA),
    plot.title = element_text(size = 10, color = "#f20707", face = "bold"),
    plot.subtitle = element_text(size = 9, color = "grey30"),
    plot.caption = element_text(size = 7, color = "grey30"),
    strip.text = element_blank()
  ) +
  labs(
    title = "Lenin & the rise and fall of political parties",
    subtitle = "Number of mentions of select political parties",
    caption = "Red bar: 1917 February/October Revolutions.\nIncludes synonyms (e.g., Bolshevism)."
  ) +
  scale_y_continuous(expand = c(0.0, 0., 0.1, 0.1)) +
  scale_x_continuous(breaks = seq(1895, 1920, 5)) +
  scale_fill_manual(
    values = c("grey60", "#f75d6f")
  )


```
```{r eval=FALSE}
parties %>%
  group_by(title, word) %>%
  summarize(n = sum(mentions))%>%
  arrange(desc(n))

parties %>%
  mutate(abbrev = gsub(".*[0-9]/(.*)/.*", "\\1", url)) %>%
  filter(!abbrev %in% tolower(month.abb)) %>%
  group_by(abbrev, word) %>%
  summarize(n = sum(mentions))%>%
  arrange(desc(n))

people %>%
  mutate(abbrev = gsub(".*[0-9]/(.*)/.*", "\\1", url)) %>%
  filter(!abbrev %in% tolower(month.abb)) %>%
  group_by(abbrev, word, title) %>%
  summarize(n = sum(mentions))%>%
  arrange(desc(n))
```
